Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-10-22 demystify.tex
updates

\begin{commentbox}[Vu:]
  I often read SOP and LORs first (\autoref{sec:ievaluate}). If I am
  persuaded by then, I would skim over other factors and advocate for
  admission (unless I see red flags in other parts). However, if I am not
  convinced, then I will likely recommend rejection (unless I see
  something stand out in other parts).
  \tcblower
  Do careful research on professors, don't mention \emph{emeritus} or adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
  Also, be careful not to send statements to the wrong schools or mix
  facts (e.g., talking about school X but mentioning working with
  profs. at school Y; and do not talk about George Washington when applying to George Mason). I have seen such statements more times than I should.
\end{commentbox}

% \subsection{Dos and Don'ts of SOP}
% \paragraph{Do's}
% \begin{enumerate}
%     \item \textbf{Be Specific}: Clearly articulate your research interests and why you are drawn to the program.
%     \item \textbf{Show Enthusiasm}: Demonstrate genuine passion for your field of study.
%     \item \textbf{Proofread}: Ensure your statement is free of grammatical and typographical errors.
%     \item \textbf{Be Personal}: Include relevant personal stories that highlight your journey and motivations.
%     \item \textbf{Tailor the Statement}: Customize your statement for each program you apply to.
% \end{enumerate}

% \paragraph{Don'ts}
% \begin{enumerate}
%     \item \textbf{Avoid Generic Statements}: Don’t use vague or general phrases.
%     \item \textbf{Don't Repeat Your Resume}: Provide new insights or context beyond what’s listed in your CV.
%     \item \textbf{Steer Clear of Jargon}: Avoid overusing technical jargon that might confuse readers.
%     \item \textbf{Don’t Exaggerate}: Avoid excessive self-praise without evidence.
%     \item \textbf{Avoid Negative Tones}: Don’t criticize past institutions or colleagues.
% \end{enumerate}
%\subsection{Common Pitfalls to Avoid}\label{sec:sop-pitfalls}

%SOP is crucial for making a strong impression on adcom reviewers. This outlines common pitfalls to avoid in your SOP.

%\begin{enumerate}
%m \textbf{Generic Statements}: Using vague or general statements that could apply to any applicant. For example, saying blank statements such as “I am passionate about computer science” without providing specific examples or evidence.
%    \item \textbf{Not Customized}: Failing to mention why you chose this univ and faculty members you want to work with. If you can submit the same SOP to multiple programs with few changes, it is not customized enough.
%    \item \textbf{Too Long}: Adcom members have to read many applications and a long SOP can be a turn-off. Keep it under 2 pages.
%    \item \textbf{Lack of Passion or Motivation}: Failing to convey genuine enthusiasm for research and the specific program.
%    \item \textbf{Grammatical and Typographical Errors}: Submitting an SOP filled with grammatical mistakes, typos, and poorly constructed sentences.
%    \item \textbf{Excessive Self-Praise}: Overly boasting about achievements without providing concrete evidence or examples to back them up.
%    \item \textbf{Irrelevant Information}: Including personal anecdotes or experiences that do not relate to the field of study or research interests.
%    \item \textbf{Unrealistic Goals}: Stating research goals that are not feasible or that show a lack of understanding of the field’s current state.
%    \item \textbf{Negative Tone}: Criticizing previous institutions, professors, or colleagues.
%    \item \textbf{Failure to Address Fit}: Not explaining why you are a good fit for the program and how the program is a good fit for your goals.
%    \item \textbf{Generic Opening and Closing Statements}: Starting or ending the SOP with clichés or unoriginal statements that do not add value.
%    \item \textbf{Lack of Customization}: Submitting the same SOP to multiple programs without tailoring it to each specific program's strengths and faculty.
%\end{enumerate}

>>>>

\begin{commentbox}[Vu:]
  I often read LORs and SOP first (\autoref{sec:ievaluate}). If I am
  persuaded by then, I would skim over other factors and advocate for
  admission (unless I see red flags in other parts). However, if I am not
  convinced, then I will likely recommend rejection (unless I see
  something stand out in other parts).
  \tcblower
  Do careful research on professors, don't mention \emph{emeritus} or adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
  Also, be careful not to send statements to the wrong schools or mix
  facts (e.g., talking about school X but mentioning working with
  profs. at school Y; and do not talk about George Washington when applying to George Mason). I have seen such statements more times than I should.
\end{commentbox}

\subsection{Kiss of Death}

\begin{itemize}
\item \textbf{Too personal} Don't talk about your personal issues, e.g., family, health, relationship, etc. This can raise concerns about your ability to handle the PhD. Don't talk about religious or political beliefs (\autoref{sec:cultural}).
\item \textbf{Criticizing} your current or previous institution or professors: just like in a job interview, don't badmouth your current or previous employer. 
\item \textbf{No concrete evidence to back up claims}: e.g., saying you are passionate about a research topic without showing any experience. This is where specific names (work submitted at conf. X), numbers (outperformed SOTA by Y \%), and examples (worked on project Z) can help. 
\item \textbf{Use flowery and AI-like language}: don't use AI to write your SOP. Not that hard to raise suspicion.
\item \textbf{Not customized to the program}: If your SOP can be sent to multiple programs with few changes, it is too generic. Do some homework and mention why you want to spend the next 7 years there.
\item \textbf{Mentioning wrong professors}: do not mention emeritus professors or those who have left.  Do your homework to see if the professors you mention are still active in research.  Teaching and adjunct faculty are often not active in mentoring PhD students (\autoref{sec:faculty-types}).
\item \textbf{Too Long and fancy format}: As mentioned, keep it under 2 pages. Don't use too much coloring or fancy fonts (like those in Words). CS academics like using LaTeX (common way to write our papers and other documents), so write your SOP using LaTeX. 
\end{itemize}
